How to improve your relationship with your future self.

### 2016-06-24 workflow.Rmd
some more again, time for lunch!

Finally, coding helps us to avoid making mistakes. For example, in our example above we may be interested in how many people protest. We may use a data file containing all protests for several years. We can then create a table with the number of protesters by copying the results manually to a document you use to write your paper. In copying we can make mistakes. So a better approach is to ask R to automatically create the table (in the format you like, with horizontal lines, 3 decimals, a title, etc) and output this in whatever file type you need (.pdf, .rtf, etc). Now when we obtain new data (for additional years, or updated figures for existing data) a new table can be created quickly, so we make less mistakes and save time!

@@

Data analysis involves a long series of decisions. Each decision
requires a justification.  Some decisions will be too small and
technical for inclusion in the published article itself. These need to
be documented in the code itself [@nagler1995coding]. Paragraphs
and citations in the publication will justify the most important
decisions. So, one must code to communicate with yourself and
others. There are two main ways to avoid forgetting the reasons you
did something with data: comment your code and tightly link your code
with your writing.^[One can also try the R command
  `put-it-in-my-mind(reason,importance="high")` to firmly place a
  reason for a decision into the mind of the analyst. I myself have
  not had much luck with this function.] [I FEEL THIS COMMENT IS FUNNY BUT CONFUSING FOR OUR LATIN AMERICAN READERS PROBABLY]

@@

More importantly, we and others want to know why we did what we
did. Such explanations may not be very clear if we have some pages of
printed code in one hand and a manuscript in the other. Keep in mind
the distinction between the 'source code' of a document (i.e. what
computation was required to produce it) and the visible, type-set page
image. Page images are great for reading, but not great for
reproducing or collaborating. The source code of any document
exchanged by the group must be available and executable.^[As of the 2016 version of this paper, this idea is now widespread and made much easier than before via online services for code sharing and collaboration such as [GitHub](http://github.com), [Open Science Framework](https://osf.io/) and [BitBucket](http://bitbucket.org), see below.]

@@

[MAARTEN IDEAS HERE? PLUS ANOTHER SECTION ABOUT ORGANIZATION?]

Where should one store data files? An obvious solution is always to
make sure that the data file used by a command file is in the same
directory as the command file. More elegant solutions require all
co-authors to have the same directory structure so that
`load("Data/thedata.rda")` means the same thing on all computers
used to work on the project. This kind of solution is one of the
things that formal version control systems do well (as discussed
a bit more in [@sec:vers-contr-prev]).

@@

@@

In the good old days, when we executed our LISREL code in batch mode,
we had no choice but to tell the machine clearly, in a few easy to
understand and informative lines, what files (with filenames no longer
than 8 characters) to use. See for example, this extremely clear code snippet:

```
DA NI=19 NO=199 MA=CM
LA=basic.lab
CM FI=basic.cov
```

The fact that we need to articulate this idea at all arises because of
graphical user interfaces: it is all too easy to use the mouse to load
a data file into memory and then to write a script to analyze this
file without ever noting the actual name or location of the data file.

@@

Jake uses github within his research group, and used it for all of
his own projects. MORE ON GITHUB
You may even obtain a two year unlimited repositories with an academic affiliation.

@@

There are some noteworthy new developments also. For example, and similar to R Markdown/KnitR in R Studio, there is Markdoc for STATA users (https://github.com/haghish/MarkDoc). People at UC Berkeley have been developing a is a web app, Jupyter Notebook (http://jupyter.org/), that enables you to make and share documents that include text, code, equations and visualizations in one INCLUDE REF TO GITHUB - USING PYTHON. For researchers working across different platforms, there is Docker (https://www.docker.com/). There are also are great notes programs that help you communicate with collaborators. For example, Simplenote is very basic but allows you to share and change notes.

>>>>

Coding helps us to avoid making mistakes. For example, in our example above we may be interested in how many people protest. We may use a data file containing all protests for several years. We can then create a table with the number of protesters by copying the results manually to a document you use to write your paper. In copying we can make mistakes. So a better approach is to ask R to automatically create the table (in the format you like, with horizontal lines, 3 decimals, a title, etc) and output this in whatever file type you need (.pdf, .rtf, etc). Now when we obtain new data (for additional years, or updated figures for existing data) a new table can be created quickly, so we make less mistakes and save time!

@@

Data analysis involves a long series of decisions. Each decision
requires a justification.  Some decisions will be too small and
technical for inclusion in the published article itself. These need to
be documented in the code itself [@nagler1995coding]. Paragraphs
and citations in the publication will justify the most important
decisions. So, one must code to communicate with yourself and
others. There are two main ways to avoid forgetting the reasons you
did something with data: comment your code and tightly link your code
with your writing.

@@

More importantly, we and others want to know why we did what we
did. Such explanations may not be very clear if we have some pages of
printed code in one hand and a manuscript in the other. Keep in mind
the distinction between the 'source code' of a document (i.e. what
computation was required to produce it) and the visible, type-set page
image. Page images are great for reading, but not great for
reproducing or collaborating. The source code of any document
exchanged by the group must be available and executable.^[As of the 2016 version of this paper, this idea is now widespread and made much easier than before via online services for code sharing and collaboration such as [GitHub](http://github.com), [Open Science Framework](https://osf.io/) and [BitBucket](http://bitbucket.org), see more below.]

@@

Where should one store data files? An obvious solution is always to
make sure that the data file used by a command file is in the same
directory as the command file. More elegant solutions require all
co-authors to have the same directory structure so that
`load("Data/thedata.rda")` means the same thing on all computers
used to work on the project. This kind of solution is one of the
things that formal version control systems do well (as discussed
a bit more below in [@sec:vers-contr-prev]).

@@

Best practice is to create a file system where you separate folders by function and separate input from output files. For example, below we show a folder structure for a paper where we look at inequality. The paper is written in .tex, data analysis in STATA (that Maarten prefers, so we see .do and .dta file extensions) in R (the program of choice for Jake, see the .R extensions):

<!-- JAKE please check if I made no mistake, also there may be a prettier way to show a folder structure (a screen shot?),feel free to change-->
00_archive
01_paper
  20160622_inequality.tex
02_data
  00_archive
  01_rawdata
  02_cleandata
  20160622_analysis.dta
  20160622_analysis.rda
03_analysis
  00_archive
  01_temp
  02_output
    01_tables
    02_figures
  20160622_analysis.do
  20160622_analysis.R
  20160622_prep_data.do
  20160622_prep_data.R

There are a couple of things to note. See that files start with a number to give them an ordering. Note also we create an 00_archive folder, here you can store older files without having to delete them permanently and keeping the number of files in your main folders to a minimum. Also there is a clear separation between raw data (data that came straight from the field, downloaded data, someone else's replication files, etc) and clean data (data ready to use). Note that in the analysis folder we created two files, one for cleaning and merging data called 20160622_prep_data and another for running the analysis called 20160622_analysis. See also the output folders, with subfolders for graphs and tables. These are place holder for R or STATA to output the results from the analysis and keeps things organized. We discuss integrated programs (that combine code and text below).

@@

@@

Jake uses github within his research group, and used it for all of
his own projects. <!-- JAKE do you want to add something else here? -->
You may even obtain a two year unlimited repositories with an academic affiliation.

@@

There are some noteworthy new developments also. For example, and similar to R Markdown/KnitR in R Studio, there is Markdoc for STATA users (https://github.com/haghish/MarkDoc). People at UC Berkeley have been developing a is a web app, Jupyter Notebook (http://jupyter.org/), that enables you to make and share documents that include text, code, equations and visualizations in one INCLUDE REF TO GITHUB - USING PYTHON. For researchers working across different platforms, there is Docker (https://www.docker.com/).

There are also are great notes programs that help you communicate with collaborators. Remember, while email is fantastic for fast communication, it is not a tool designed for project management. It may work for small projects with few co-authors but when the number of collaborators increase it becomes very difficult to keep track. There are some great online systems for task management, most of which have mobile apps also, examples are Flow (www.getflow.com), Asana (www.asana.com), Wrike (www.wrike.com) <!-- JAKE do you know others? --> . There also are great programs for notes. For example, Simplenote is very basic but allows you to share and edit notes. Others are Evernote (www.evernote.com) and for Windows users there is OneNote.
